---
title: "CPSC-375 Homework 4"
author: "Kenn Son, Hamid Suda, Vivian Truong"
date: "2/23/2022"
output:
  pdf_document: default
  html_document: default
---
```{r install}
#install.packages("nycflights13")
library(nycflights13)
library(tidyverse)
flights
```

1. List data only for flights that departed on February 12, 2013.

```{r q1}
flights %>% filter(year=="2013", month=="2", day=="12")
```

2. List data only for flights that were delayed (both arrival and departure) by more than 2 hours.

```{r q2}
flights %>% filter(dep_delay > 200) %>% filter(arr_delay > 200)
```

3. List data only for flights that were delayed (either arrival or departure) by more than 2 hours.
```{r q3}
flights %>% filter(dep_delay > 200|arr_delay > 200)
```

4. List data only for flights that were operated by United, American, or Delta.
```{r q4}
flights %>% filter(carrier == "UA" | carrier == "AA" | carrier == "DL")
```

5. Sort data in order of fastest flights (air_time).
```{r q5}
flights %>% arrange(air_time)
```

6. Sort data in order of longest duration flights (air_time).
```{r q6}
flights %>% arrange(desc(air_time))
```

7. Show only the origin and destination of flights sorted by longest flights.
```{r q7}
flights %>% arrange(desc(air_time)) %>% select(origin, dest)
```

8. Add a new variable that indicates the total delay (both departure and arrival delay).
```{r q8}
flights %>% mutate(total_delay = flights$dep_delay + flights$arr_delay)
```

9. Show only the origin and destination of flights sorted by descending order of total delay.
```{r q9}
flights %>% mutate(total_delay = flights$dep_delay + flights$arr_delay) %>% 
  arrange(desc(total_delay)) %>% select(origin,dest)
```

10. Show only the origin and destination of 10 most delayed flights [Hint: there are multiple ways of solving this. Some additional functions that you will find useful are head(), slice(), min_rank().]
```{r q10}
flights %>% mutate(total_delay = flights$dep_delay + flights$arr_delay) %>% 
  arrange(desc(total_delay)) %>% top_n(10, total_delay)
```

11. Show the average total delay for all flights
```{r q11}
flights %>% mutate(total_delay = flights$dep_delay + flights$arr_delay) %>% 
  summarise(mean(total_delay, na.rm = TRUE))
```

12. Show the average total delay for every departure city.
```{r q12}
flights %>% mutate(total_delay = flights$dep_delay + flights$arr_delay) %>% 
  group_by(origin) %>% summarise(mean(total_delay, na.rm = TRUE))
```

13. Show the average total delay for every departure-arrival city pair.
```{r q13}
flights %>% mutate(total_delay = flights$dep_delay + flights$arr_delay) %>% 
  group_by(origin, dest) %>% summarise(mean(total_delay, na.rm = TRUE))
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

