---
title: "CPSC375Homework 5"
author: "Kenn Son, Hamid Suda, Vivian Truong"
date: "3/8/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

1. Consider the two tables shown below called population and countyseats.

population:
```{r pop1}
state <- c("California", "California", "California", "California")
county <- c("Orange", "Orange", "Los Angeles", "Los Angeles")
year <- c(2000,2010,2000,2010)
pop <- c(2846289,3010232,3694820,3792621)
population <- data.frame(state, county, year, "Population" = pop)
x <- as_tibble(population)
x
```

countyseats:
```{r pop2}
statename <- c("California", "California", "California", "Oregon")
countyname <- c("Orange", "Los Angeles", "San Diego", "Wasco")
countyseat <- c("Santa Ana", "Los Angeles", "San Diego", "The Dalles")
countyseats <- data.frame(statename, countyname, countyseat)
y <- as_tibble(countyseats)
y
```

You should be able to calculate the output by hand though you may use R to check your answer.
Draw the output table from the following operations (you should be able to calculate the output by hand though you may use R to check your answers).

|       a) population %>% inner_join(countyseats)
|           - Error since we don't know what is being compared
|       b) population %>% inner_join(countyseats, by=c(state="statename"))

| state | county |year|population|countyname|countyseat|
|:-----:|:------:|:--:|:--------:|:--------:|:--------:|
|California|Orange|2000|2846289|Orange|Santa Ana|
|California|Orange|       2000|    2846289| Los Angeles| Los Angeles|
|California|Orange|       2000|    2846289| San Diego |  San Diego  |
|California|Orange|       2010|    3010232| Orange     | Santa Ana  |
|California| Orange|       2010|    3010232| Los Angeles| Los Angeles|
|California| Orange|       2010|    3010232| San Diego  | San Diego  |
|California| Los Angeles|  2000|    3694820| Orange   |   Santa Ana  |
|California| Los Angeles|  2000|    3694820| Los Angeles| Los Angeles|
|California| Los Angeles|  2000|    3694820| San Diego |  San Diego  |
|California| Los Angeles|  2010|    3792621| Orange    |  Santa Ana  |
|California| Los Angeles|  2010|    3792621| Los Angeles |Los Angeles|
|California| Los Angeles|  2010|    3792621| San Diego  | San Diego|

|       c) population %>% inner_join(countyseats, by=c(state="statename",  county="countyname"))
```{r test, include=FALSE}
x %>% inner_join(y, by=c(state="statename"))
```

| state | county |year|population| countyseat |
|:-----:|:------:|:--:|:--------:|:----------:|
| California | Orange       | 2000    | 2846289 | Santa Ana   |
| California | Orange       | 2010    | 3010232 | Santa Ana   |
| California | Los Angeles  | 2000    | 3694820 | Los Angeles |
| California | Los Angeles  | 2010    | 3792621 | Los Angeles |


|       d) population %>% inner_join(countyseats, by=c(state="statename",  county="countyname", year="countyseat"))

| state | county |year|population| countyseat |
|:-----:|:------:|:--:|:--------:|:----------:|
| California | Orange       | 2000    | 2846289 | Santa Ana   |
| California | Orange       | 2010    | 3010232 | Santa Ana   |
| California | Los Angeles  | 2000    | 3694820 | Los Angeles |
| California | Los Angeles  | 2010    | 3792621 | Los Angeles |

|   2. Consider the billboard dataset that is supplied with the tidyverse which shows the Billboard top 100 song rankings in the year 2000. Apply the tidyverse’s data wrangling verbs to answer these questions. For each question, give only the code.

|       a) Show for each track, how many weeks it spent on the chart

```{r 2a}
billboard %>% select(-artist) %>% select(-date.entered) %>%
  pivot_longer(-track, names_to = 'Week',values_to = 'Place',values_drop_na = TRUE) %>%
  group_by(track) %>% summarize('Count'=n())
```

|       b)List tracks in decreasing order of number of weeks spent on the chart

```{r 2b}
billboard %>% select(-artist) %>% select(-date.entered) %>%
  pivot_longer(-track, names_to = 'Week',values_to = 'Place',values_drop_na = TRUE) %>%
  group_by(track) %>% summarize('Count'=n()) %>% arrange(desc(Count))
```

|       c)Show for each track, its top rank

```{r 2c}
billboard %>% select(-artist) %>% select(-date.entered) %>%
  pivot_longer(-track, names_to = 'Week',values_to = 'Place',values_drop_na = TRUE) %>%
  group_by(track) %>% summarise('TopRank
                                ' = min(Place))
```

|       d)List tracks in increasing order of its top rank

```{r 2d}
billboard %>% select(-artist) %>% select(-date.entered) %>%
  pivot_longer(-track, names_to = 'Week',values_to = 'Place',values_drop_na = TRUE) %>%
  group_by(track) %>% summarise('TopRank' = min(Place)) %>%
  arrange(TopRank)
```

|       e)Show for each artist, their top rank

```{r 2e}
billboard %>% select(-track, -date.entered) %>% 
  pivot_longer(-artist, names_to = 'Week', values_to = 'Place',values_drop_na = TRUE) %>%
  group_by(artist) %>% summarise('TopRank' = min(Place))
```

|       f)List artists in increasing order of their top rank

```{r 2f}
billboard %>% select(-track, -date.entered) %>% 
  pivot_longer(-artist, names_to = 'Week', values_to = 'Place',values_drop_na = TRUE) %>%
  group_by(artist) %>% summarise('TopRank' = min(Place)) %>% arrange(TopRank)
```

|       g)List tracks that spent more than 35 weeks in the charts

```{r 2g}
billboard %>% select(-artist) %>% select(-date.entered) %>%
  pivot_longer(-track, names_to = 'Week',values_to = 'Place',values_drop_na = TRUE) %>%
  group_by(track) %>% summarize('Count'=n()) %>% filter(Count > 35)
```

|       h) List tracks that spent more than 35 weeks in the charts along with their artists

```{r 2h}
billboard %>% select(-date.entered) %>% 
  pivot_longer(
    -c(artist, track), names_to = 'Week',values_to = 'Place',values_drop_na = TRUE
    ) %>%
  group_by(track, artist) %>% summarize('Count'=n()) %>% filter(Count > 35)
```

Hint: First, convert to a tidy table. Show code first for this step. All the above questions can then be answered with a single data pipeline.

3. The demographics.csv  file (available in the Datasets module on Canvas) gives the proportion of a country’s population in different age groups and some other demographic data such as mortality rates and expected lifetime. You can read a CSV file into a tibble using tidyverse’s read_csv(), like so: demo <- read_csv("demographics.csv")

```{r inputDemo}
demo <- read_csv("demographics.csv")
```

|       a) The data is not “tidy”. In 2-3 sentences, explain why.

|           The data is not tidy because the Series Code and YR2015 are messy to read. 
|           The Series name is also not that helpful as theres its already categorized on with the code.
|           The table can group by the country names and then have multiple series code column.

|       b) Transform the table to tidy data with one country per row. [Give code]
```{r 3b}
new_demo <- demo %>% select(-`Series Name`) %>%
  pivot_wider(names_from = `Series Code`, values_from = YR2015) %>%
  group_by(`Country Name`)
new_demo
```

|       c) Add the male/female population numbers together (i.e., ignore sex-related differences). [Hint: You will have to mutate for every pair of columns, e.g., mutate(SP.POP.0014.IN=SP.POP.0014.MA.IN+SP.POP.0014.FE.IN) [Give code]
```{r 3c}
combine_demo <- new_demo %>% mutate(SP.POP.80UP=SP.POP.80UP.MA+SP.POP.80UP.FE) %>%
  mutate(SP.POP.1564=SP.POP.1564.MA.IN+SP.POP.1564.FE.IN) %>%
  mutate(SP.POP.0014.IN=SP.POP.0014.MA.IN+SP.POP.0014.FE.IN) %>%
  mutate(SP.DYN.AMRT=SP.DYN.AMRT.MA+SP.DYN.AMRT.FE) %>%
  mutate(SP.POP.TOTL.IN=SP.POP.TOTL.MA.IN+SP.POP.TOTL.FE.IN) %>%
  mutate(SP.POP.65UP.IN=SP.POP.65UP.MA.IN+SP.POP.65UP.FE.IN) %>%
  select(c(`Country Name`, `Country Code`, SP.DYN.LE00.IN, SP.URB.TOTL,
          SP.POP.80UP, SP.POP.1564,SP.POP.0014.IN,SP.DYN.AMRT,SP.POP.TOTL.IN,
          SP.POP.65UP.IN))
combine_demo
```

|       d) Write code to show the top 5 countries with the lowest proportion of the population below 14 years old (i.e., SP.POP.0014.IN/SP.POP.TOTL) [Code, and list of 5 countries]
```{r 3d}
demo.0014 <- combine_demo %>%
  mutate(`Percent of 14 years and Under` = SP.POP.0014.IN/SP.POP.TOTL.IN) %>%
  select(c(`Country Name`,`Percent of 14 years and Under`)) %>% 

    arrange(`Percent of 14 years and Under`)
demo.0014[1:5,] #couldn't figure out top_n
```
